### 2019-11-25 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3
mathematical statistics, ch3

@@

@@

gte.t <- perms[ Xd >= mean(mice.t),]
pretty_kable(gte.t, "Mean Treatment Greater than Observed")

@@

@@

@@

a.) Which P-value, 0.03 or 0.006 provides stronger evidence for the alternative hypothesis?

b.) Which P-value, 0.095 or 0.04 provides stronger evidence that chance alone might account for the observed result?

@@

@@

@@

@@

>>>>

# 5 choose 3 for treatment

@@

stopifnot(nrow(perms) == choose(5, 3))

@@

gte.t <- perms[ Xd >= mean(mice.t),]
pretty_kable(gte.t, "Mean Treatment Greater than or Equal to Observed")

@@

+ Yes, a 1/10,000 chance is pretty rare.

@@

+ A 1/10 chance is fairly common.

@@

__a.)__ Which P-value, _0.03_ or _0.006_, provides stronger evidence for the alternative hypothesis?

_0.03_ provides stronger evidence for the alternative hypothesis.

__b.)__ Which P-value, 0.095 or 0.04, provides stronger evidence that chance alone might account for the observed result?

_0.095_ provides stronger evidence that chance alone is responsible for the observed result.

@@

__Answer__: We need to account for the original observed result.

@@

Flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
                               header = T))

@@

```{r, echo = T, fig.height=3.5, fig.width=8}

Flights[, .(Delay = mean(Delay)), by = Carrier]

observed <- mean(Flights[Carrier == "UA"]$Delay) - mean(Flights[Carrier == "AA"]$Delay)

N <- 10e2 - 1
results <- numeric(N)

for(i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
   results[i] <- mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
}

p <- (sum(results[results >= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..)) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", p ))

@@

```{r, echo = T, fig.height=3.5, fig.width=8}

Flights[, .(Delay = mean(Delay)), by = Month]

observed <- mean(Flights[Month == "May"]$Delay) - mean(Flights[Month == "June"]$Delay)

N <- 10e2 - 1
results <- numeric(N)

for(i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
   results[i] = mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
}

p <- (sum(results[results <= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..)) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)))
